\documentclass[12pt]{article}
\usepackage [margin=1in]{geometry}
\usepackage{amsmath, amssymb}

\begin{document}

\noindent Isaac Kim

\noindent CS 487

\noindent Professor Baldimtsi

\noindent Mar 31, 2023

\begin{center} 
\textbf{Homework 4}
\end{center}

\noindent \textbf{1.} $MAC'(k,m)$ = \emph{MAC}$(k,m)||\emph{MAC}(k,m)$.
\\
\\
Verification works as follows: Let $t = MAC(k,m)$ and $t' = MAC'(k,m)$. Then $\emph{Vrfy}'(m, t')=1$ if and only if $t' = t||t$ and $\emph{Vrfy}(m,t)=1$
\\
\\
Contrapositive: If $MAC'$ is \emph{not} secure, then $MAC$ is \emph{not} secure. This means that there exists an adversary $A'$ such that the probability that $A'$ can successfully forge a message with a valid tag is $p(n)$, where $p(n)$ is some non-negligible value.
\\
\\
Reduction: Using $A'$, we can construct an adversary $A$ that breaks $MAC$.

Step 1. $A$ generates a tag $t$ and makes $t'$ such that $t' = t||t$ and $|t'| =2t$

Step 2. $A$ runs $A'$ and gives $t' = t||t$ to $A'$

Step 3. $A'$ can then successfully forge a valid message $m$ for $t'$ with probability $p(n)$ such that $\emph{Vrfy}'(m, t')=1$

Step 4. In doing so, $A'$ forges $t' = MAC'(k,m) = MAC(k,m)||MAC(k,m) = t||t$

Step 5. $A$ can then output $(m, t)$. The probability that $\emph{Vrfy}(m,t)=1$ is equal to $p(n)$
\\
\\
Hence, $A$ is a good adversary for $MAC$.
\\
\\
\noindent \textbf{2.1)} Let $m_0 = x_0 \oplus w_0$, $m_1 = x_1 \oplus w_1$, $m_0 = 000, m_1=0000$. Then, it must follow that $x_0 = w_0$ and $x_1 = w_1$. Note that $m_0 \neq m_1$. Then, $H(m_0) = H(x_0) \oplus H(w_0) = 0^n = H(x_1) \oplus H(w_1) = H(m_1)$. Then, $H(m_0) = H(m_1)$, but $m_0 \neq m_1$. Hence, $H$ is not collision resistant.
\\
\\
\noindent \textbf{2.2)} Let $x_\alpha = x_1 || x_2$, and $x_\beta = x_2 || x_1$, $x_1 \neq x_2$. Note that $x_\alpha \neq x_\beta$. Then, $H_s^a(x_\alpha) = H_s(x_1) \oplus H_s(x_2) = H_s(x_2) \oplus H_s(x_1) = H_s^a(x_\beta)$. Then, $H_s^a(x_\alpha)= H_s^a(x_\beta)$ but $x_\alpha \neq x_\beta$. Hence, $H_s^a$ is not collision resistant.
\\
\\
\noindent \textbf{3)} $H_s^b(x) = H_s^1(x)||H_s^2(x)||H_s^3(x)$.
\\
\\
Consider the case where only $H_s^1$ is collision resistant. Then, $H_s^b(x)$ is also collision resistant. To show this, we will show equivalently by contrapositive that if $H_s^b$ is \emph{not} collision resistant, then $H_s^1(x)$ is \emph{not} collision resistant.
\\
\\
This implies that there exists an adversary $A_b$ for $H_s^b$ such that the probability that $A_b$ can output $x, x', x \neq x'$ with $H_s^b(x) = H_s^b(x')$ is $p(n)$, where $p(n)$ is a non-negligible value.
\\
\\
Let $A_1$ be an adversary for $H_s^1$. $A_1$ runs $A_b$, and $A_b$ outputs $x, x', x \neq x'$. Since $A_b$ is a good adversary for $H_s^b$, with probability $p(n)$, it holds that $H_s^b(x) = H_s^1(x)||H_s^2(x)||H_s^3(x) =  H_s^1(x')||H_s^2(x')||H_s^3(x') = H_s^b(x')$. This implies that $H_s^1(x) = H_s^1(x')$. Then, $A_1$ can output $x,x',x \neq x'$ since $H_s^1(x) = H_s^1(x')$. The probability of success for $A_1$ is the same as the probability of success for $A_b$, namely $p(n)$. Hence, $A_1$ is a good adversary for $H_s^1$. The proof for cases where $H_s^2$ or $H_s^3 $ are the only collision resistant functions is similar.
\\
\\
\noindent \textbf{4)} 

\noindent \textbf{a)} If $H_s$ is collision resistant, then $H_s'$ is collision resistant. Contrapositive: If $H_s'$ is \emph{not} collision resistant, then $H_s$ is \emph{not} collision resistant.
\\
\\
Let $A'$ be an adversary for $H_s'$ that can find a collision for $y = H_s'(x_1, x_2,...,x_{2^h})$. Let $a, b$ be the children of the root node $y$ and $a', b'$ be the children of the root node $y'$, $a \neq a', b \neq b'$. 
\\
\\
Let $A'$ find a collision $x, x'$ such that $x = x_1, x_2,...,x_{2^h}, x' = x_1', x_2',...,x_{2^h}', x \neq x'$. Note that $y = H_s'(x_1, x_2,...,x_{2^h}) = H_s(a||b)$ and $y' = H_s'(x_1', x_2',...,x_{2^h}') = H_s(a'||b')$. Since this is a collision, 
\begin{center}
$y = H_s'(x) = H_s'(x_1, x_2,...,x_{2^h}) =  H_s'(x_1', x_2',...,x_{2^h}')= H_s'(x')=y' \rightarrow  H_s(a||b)=H_s(a'||b')$. 
\end{center}
Hence, $A'$ has also found a collision for $H_s$.
\\
\\
\noindent \textbf{b)} Consider $y_1 = H_s'(x_1, x_2) = H(x_1||x_2), y_2 =H_s'(x_1, x_2,x_1,x_2) = H(H(x_1||x_2)||H(x_1||x_2)), y_3 = H_s'(y_1, y_1)$. Then, for $y_1, h_1 = 1$, for $y_2, h_2 = 2$ and for $y_3, h_3 = 1$. Then, $y_2 =H_s'(x_1, x_2,x_1,x_2) = H(H(x_1||x_2)||H(x_1||x_2)) = H_s'(y_1, y_1) = y_3$, but $h_2 = 2 \neq 1 = h_3$. Hence, if $h$ is not fixed, this construction is not secure.


\end{document}