\documentclass[12pt]{article}
\usepackage [margin=1in]{geometry}
\usepackage{amsmath, amssymb}

\begin{document}

\noindent Isaac Kim

\noindent CS 487

\noindent Professor Baldimtsi

\noindent Mar 1, 2023

\begin{center} 
\textbf{Homework 3}
\end{center}

\noindent \textbf{1.1)} Consider an adversary \emph{A} that initially queries a message $m_\alpha = <m_{\alpha_1},...,m_{\alpha_t}>$ to the Challenger, and gets a ciphertext $c_\alpha = <c_{\alpha_0},c_{\alpha_1},...,m_{\alpha_t}>$ in return, where $c_{\alpha_0} = IV$. By Kerckhoff's Principle, \emph{A} knows that \emph{S} sets $IV_i = IV_{i-1} + 1$. Let $IV_{i-1}$ be the $IV$ for $m_\alpha$. Then, since $m_\alpha$ was just queried, \emph{A} can pick messages $m_0, m_1$ such that for the first block of $m_0$:
\begin{center}
$m_{0_1} \oplus IV_i = m_{\alpha_1} \oplus IV_{i-1}$

$m_{0_1} \oplus IV_i = c_{\alpha_1}$

$m_{0_1} = c_{\alpha_1} \oplus IV_i$
\end{center}
Then, pick $m_1$ such that the first block of $m_1$, or $m_{1_1} \neq m_{0_1}$. The following blocks of $m_0, m_1$ can be arbitrary. Then, when \emph{A} sends $m_0, m_1$ to the Challenger and gets back $c*=<c*_0,c*_1,...,c*_t>$ where $c*_0=IV_i$, \emph{A} only needs to compare the second blocks of $c*$ and $c_\alpha$. If $c*_1 = c_{\alpha_1}$, then \emph{A} outputs $b' = 0$, and otherwise, outputs $b' = 1$. Then, Pr[$A$ wins] $= 1 = \frac{1}{2} + \frac{1}{2} =\frac{1}{2} + p(n)$ where $p(n)$ is a non-negligible value. Hence, this scheme is \emph{not} CPA secure.
\\
\\
\noindent \textbf{1.2)} Let \emph{A} be an adversary that queries messages $m_0 = 0^n$, $m_1=1^n$ to the Challenger. The Challenger will then flip a bit $b \in \{0,1\}$, encrypt $c* =$ Enc$(k, m_b)$ for some key $k$, and return $c* = <c*_0, c*_1,...c*_t>$ to \emph{A}. 
\\
\\
\noindent Note that OFB Decryption is as follows: For some ciphertext $c = <c_0, c_1,...c_t>$, $c_0 = IV$, $m_1 = F_k(c_0)\oplus c_1$,... in general, $m_n = F_k(F_k(c_{n-1})) \oplus c_n$. 
\\
\\
\noindent Then, \emph{A} can query $s$ to the decryption oracle, where $s = <c*_0, \bar{c*_1},...c*_t>$, $\bar{c*_1} = c*_1$ with its first bit flipped.
Since $s \neq c*$, this is allowed, and $A$ will get back a message $m_s=$ $10^n$ or $01^n$. If the $m_s = 10^n$, then $A$ outputs $b'=0$, and otherwise, $m_s = 01^n$ and $A$ outputs $b' = 1$. Then, Pr[$A$ wins] $= 1 = \frac{1}{2} + \frac{1}{2} =\frac{1}{2} + p(n)$ where $p(n)$ is a non-negligible value. Hence, this scheme is \emph{not} CCA secure.
\\
\\

\noindent \textbf{2.1)} Let $A$ be an adversary that queries message $m_\beta = m_1||...||m_l$, $|m_\beta| = ln$, to the MAC oracle Mac$_k(\cdot)$, and gets back $t_\beta \leftarrow $Mac$_k(m_\beta)$. By Kerckhoff's Principle, $A$ knows that $t_\beta = F_k(m_1) \oplus ... \oplus F_k(m_l)$. Then, $A$ can output $(m_\alpha, t_\beta)$, where $m_\alpha = F_k(m_l) \oplus ... \oplus F_k(m_1)$. Then, Vrfy$(m_\alpha, t_\beta) = 1$ since $t_\beta$ is a valid tag for $m_\alpha$, but $m_\alpha$ was never previously authenticated. Hence, our MAC scheme is not secure.
\\
\\
\noindent \textbf{2.2)} Let $m_\alpha = m_3||m_1$, $m_\beta = m_1||m_2$, $|m_\alpha| = |m_\beta| = 2n$ and $|m_1|=|m_2|=|m_3|=n$. Let $A$ be an adversary that queries messages $m_\alpha, m_\beta$ to the MAC oracle Mac$_k(\cdot)$, and gets back $t_\alpha \leftarrow $Mac$_k(m_\alpha)$ and $t_\beta \leftarrow $Mac$_k(m_\beta)$. By Kerckhoff's Principle, $A$ knows that $t_\alpha = F_k(m_3)||F_k(F_k(m_1))$ and $t_\beta = F_k(m_1)||F_k(F_k(m_2))$. Then, $A$ can output $(m_\gamma, t_\beta)$, where $m_\gamma = m_1||m_1$ and $t_\gamma = F_k(m_1)||F_k(F_k(m_1))$. Then, Vrfy$(m_\gamma, t_\gamma) = 1$ since $t_\gamma$ is a valid tag for $m_\gamma$, but $m_\gamma$ was never previously authenticated. Hence, our MAC scheme is not secure.
\end{document}